### 2025-01-15 analysis/Supplementary/supplementary_tables.qmd
Supplementary tables improved and organised. WIP render.

---
title: "SOM result tables"
author:
  - Jane Doe:
      correspondence: "yes"
      email: janedoe@fosg.org
      orcid: 0000-0003-1689-0557
      institute:
        - fosg
        - fop
  - John Q. Doe:
      institute: fosg
      orcid: 0000-0003-1689-0558
  - Peder Ås:
      institute: fosg
      orcid: 0000-0003-1689-0559
  - Juan Pérez:
      orcid: 0000-0003-1689-0551
      institute:
        - name: Acme Corporation
  - Max Mustermann:
      orcid: 0000-0003-1689-0552
institute:
  - fosg:
      name: Formatting Open Science Group
      address: 23 Science Street, Eureka, Mississippi, USA
  - fop: Federation of Planets
title-block-published: "Last updated"  
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: false
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: ../paper/references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
---

```{r}
#| label: library

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(knitr)

```

```{r}
#| label: get-data

# Technological data
gravettian_tech <- read.csv(here::here('analysis/data/raw_data/VBT_6_7_JB.csv'))
proto_tech <- read.csv(here::here('analysis/data/raw_data/VBT_4E_5_JB.csv'))
solutrean_JC <- read.csv(here::here('analysis/data/raw_data/VBR_ABC_JC.csv'))
solutrean_points <- read.csv(here::here('analysis/data/raw_data/shelter_points_JC.csv'))

# RM data
shelter_RM <- read_csv(here::here('analysis/data/raw_data/MACRO_IND_SHELTER.csv')) # Shelter RM analysis
terrace_RM <- read_csv(here::here('analysis/data/raw_data/MACRO_IND_TERRACE.csv')) # Terrace RM analysis

# Context data
context <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_context.csv')) # Terrace Field DB - lithic artefact info
xyz <- read_csv(here::here('analysis/data/raw_data/VB_Terrace_xyz.csv')) # Terrace Field DB - just coordinates

```

```{r}
#| label: databases-setup

# Terrace dataset setup
cols_to_concat_terrace <- c("UNIT", "ID")
context$ID <- str_squish(context$ID) # Remove empty spaces from ID field
xyz$ID <- str_squish(xyz$ID) # Remove empty spaces from UNIT field

proto_tech_WIP <- proto_tech %>%
  tidyr::unite(col='ID', all_of(cols_to_concat_terrace), sep="-", remove=TRUE) %>% 
  rename(CLASS = Class, COREPREPPROD = CorePreparProd, PIECECOMPLETENESS = PieceCompleteness, CORTEX = Cortex, CORTEXLOCATION = CortexLocation, PLATFORMTYPE = PlatformType, PLATFORMCORTEX = PlatformCortex, SCARCOUNT = ScarCount, THICKNESS = Thickness,
        MAXWIDTH = MaxWidth, LENGTH = Length, COREPLATFORM = CorePlatform, MAINFACECORTEX = MainFaceCortex, 
        MAINFACESCARCOUNT = MainFaceScarCount, COREABANDON = CoreAbandon, RETOUCHEDPIECETYPOLOGY = RetouchedPieceTypology) %>% 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Right truncation on retouched piece" ~ "Truncation",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" | RETOUCHEDPIECETYPOLOGY == "Backed bladelet parcial" ~ "Backed bladelet",
    RETOUCHEDPIECETYPOLOGY == "Dentilhado" ~ "Denticulate",
    RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) %>% 
  dplyr::select(ID, CLASS, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE, PLATFORMCORTEX,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, MAINFACECORTEX, MAINFACESCARCOUNT, COREABANDON, RETOUCHEDPIECETYPOLOGY)

proto_original_chert <- proto_tech %>% 
  filter(RawMaterial %in% c("Chert", "Chalcedony"))

gravettian_tech <- gravettian_tech %>% 
  mutate(RETOUCHEDPIECETYPOLOGY = case_when(
    RETOUCHEDPIECETYPOLOGY == "Splintered piece with notch" | RETOUCHEDPIECETYPOLOGY == "Splintered piece with lateral retouch" ~ "Splintered piece",
    RETOUCHEDPIECETYPOLOGY == "Retouched flake siret fragment" | RETOUCHEDPIECETYPOLOGY == "Retouched flake frag" ~ "Retouched flake fragment",
    RETOUCHEDPIECETYPOLOGY == "Retouched bladelet frag" ~ "Retouched bladelet fragment",
    RETOUCHEDPIECETYPOLOGY == "Multiple burin" | RETOUCHEDPIECETYPOLOGY == "Burin splintered piece" | RETOUCHEDPIECETYPOLOGY == "Burin on truncation" ~ "Burin",
    RETOUCHEDPIECETYPOLOGY == "Double backed bladelet" ~ "Backed bladelet",
    TRUE ~ RETOUCHEDPIECETYPOLOGY)) %>% 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    TRUE ~ CLASS))

terrace_tech <- bind_rows(proto_tech_WIP, gravettian_tech)

terrace_RM_WIP <- terrace_RM %>%
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "T2",
    RMTYPE == "Chalcedony" ~ "T1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "T3",
    RMTYPE == "TYPE 7" ~ "T4",
    RMTYPE == "TYPE 12" ~ "T5",
    RMTYPE == "TYPE 11" ~ "T6",
    RMTYPE == "TYPE 13" ~ "T7",
    RMTYPE == "TYPE 11F" ~ "T8",
    RMTYPE == "TYPE 13B" ~ "T9",
    RMTYPE == "Oolitic" ~ "T10",
    RMTYPE == "TYPE GB" ~ "T11",
    .default = RMTYPE)) %>% 
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%
  distinct(ID, .keep_all = TRUE)

context <- context %>%
  dplyr::rename("UNIT" = "Unit") %>% # Renaming first column from Unit to UNIT to use cols_to_concat
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) %>% # Unite UNIT with ID using cols_to_concat
  select(ID, Level, Spit, Code) %>% # Selecting important variables: ID refers to artifact ID; Level refers to archaeologycal layer; Spit refers to artificial levels within an archaeological layer; Code refers to type of artifact (e.g., LITHIC or BONE)
  distinct(ID, .keep_all = TRUE)

xyz <- xyz %>%
  unite(col='ID', cols_to_concat_terrace, sep="-", remove=FALSE) %>% # Unite UNIT with ID using cols_to_concat
  filter(Suffix == 0) %>% # To avoid several entries related to several points of the same artefact
  select(ID, X, Y, Z) %>% # Selecting important variables
  distinct(ID, .keep_all = TRUE)

field_data <- full_join(xyz, context, by = "ID") # Join field tables

terrace_tech_WIP <- left_join(terrace_tech, field_data, by = "ID") # Join field and RM type tables
terrace_dataset <- full_join(terrace_tech_WIP, terrace_RM_WIP, by = "ID") # Base 12690
terrace_dataset <- terrace_dataset %>%
  filter(!Level %in% c("3", "4", "4C", "4D")) %>%  # Dropping NAs on the technological class drops dataset to 1651
  drop_na(RMTYPE) %>%  # Dropping NAs on the RM database drops dataset to 3635
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 1647

# Mutate level 5 to add phases: top of 5 (4E/5) and bottom of 5 (5), following Belmiro et al. (2020).
phase_dataset <- terrace_dataset |>
  mutate(Level = case_when(
    Level == 5 & Z > 24.1 ~ "4E/5",
    Level == 5 & Z < 24.1 ~ "5",
    Level == "7C" | Level == 8 ~ "7", # Collapsing samples from Level 7C and Level 8 into 7 to homogenise dataset
    TRUE ~ Level)) |> # Adding technocomplex name based on the levels
  mutate(Technocomplex = case_when(
    Level == "4E" | Level == "4E/5" | Level == 5 ~ "Proto-Solutrean",
    Level == 6 | Level == 7 ~ "Gravettian"))

# Shelter dataset setup
cols_to_concat_shelter <- c("YEAR", "ID") # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

solutrean_WIP <- solutrean_JC %>% 
  filter(Layer %in% c("A", "B", "C")) %>% # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony")) %>% # Filtering raw materials
  filter(Screen != "Crivo") %>%  # Filtering out all bucket lithics since they were not included in the RM analysis
  mutate_at("Date", str_replace, "20", "") %>% # Removing initial part of the year to match IDs
  dplyr::rename("YEAR" = "Date") %>% 
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=FALSE) # Concatenate YEAR and ID variables in one single variable (e.g. 05-500)

# Dataset filtering for chert/chalcedony and levels (A-C), 3298 + 61 solutrean points, n=3359
solutrean_original_chert <- solutrean_JC %>% 
  filter(Layer %in% c("A", "B", "C")) %>% # Filtering Solutrean layers
  filter(Raw.material %in% c("Chert", "Chalcedony"))

# In Cascalheira 2010 Retouched piece was not a category but a variable. To match the other DBs, this category was added
solutrean_WIP <- solutrean_WIP %>% 
  mutate(Class = case_when( # This alteration is being done on the original class column
    Retouched == "Yes" ~ "Retouched piece", # Whenever retouch is present, class is now Retouched piece
    TRUE ~ Class)) %>% 
  mutate(CLASS = case_when( # Creating a column for clustered classes, to match the other DBs
    Class == "Blade" | Class == "Bladelet" ~ "ElongBlank", # Cluster blades with bladelets. The distinction can come later through a mutate and the measurements
    Class == "Core front" | Class == "Core tablet" | Class == "Core trimming elem." | Class == "Crested piece" ~ "CorePrepProd", # Cluster all preparation products
    Class == "Fragment" ~ "Shatter",
    Class == "Flake distal frag." | Class == "Flake mesial frag." | Class == "Flake prox. frag." ~ "FlakeFrag", # Cluster flake frags into a single category
    Class == "Blade distal frag." | Class == "Blade mesial frag." | Class == "Blade prox. frag." 
    | Class == "Bladelet distal frag." | Class == "Bladelet mesial frag." | Class == "Bladelet prox. frag." ~ "ElongBlankFrag", # Cluster elong frags into a single category
    TRUE ~ Class)) %>% 
  mutate(COREPREPPROD = case_when( # Creating a column for the types of core preparation and maintenance products (COREPREPPROD)
    Class == "Core front" ~ "Core front",
    Class == "Core tablet" ~ "Core tablet",
    Class == "Core trimming elem." ~ "Core trim",
    Class == "Crested piece" ~ "Crested piece")) %>% 
  mutate(PIECECOMPLETENESS = case_when( # Creating a column for the portions present in flake or elongblank fragments to match other DBs
    Class == "Flake distal frag." | Class == "Blade distal frag." | Class == "Bladelet distal frag." ~ "Distal",
    Class == "Flake mesial frag." | Class == "Blade mesial frag." | Class == "Bladelet mesial frag." ~ "Mesial",
    Class == "Flake prox. frag." | Class == "Blade prox. frag." | Class == "Bladelet prox. frag." ~ "Proximal")) %>% 
  rename(PLATFORMTYPE = Platform_type) %>% # Renaming column
  mutate(PLATFORMTYPE = case_when( # Renaming variables to match correspondent variables in the other DBs
    PLATFORMTYPE == "Multifacetted" ~ "Faceted",
    PLATFORMTYPE == "Unfacetted" ~ "Plain",
    TRUE ~ PLATFORMTYPE)) %>%
  mutate(Type = case_when( # Translating retouched types from Zilhão 1997 to simplified typologies
    Type == "92b" ~ "Retouched piece fragment",
    Type == "92a" ~ "Retouched blank",
    Type == "76" ~ "Splintered piece",
    Type == "75" ~ "Denticulate",
    Type == "74" ~ "Notch",
    Type == "66" ~ "Retouched blade",
    Type == "61" ~ "Truncation",
    Type == "28" ~ "Burin",
    Type == "7" | Type == "5b" | Type == "4"  ~ "Endscrapper",
    TRUE ~ Type)) %>%
  rename(LEVEL = Layer, UNIT = Unit, CORTEX = Percent_of_cortex, CORTEXLOCATION = Loc_of_cortex, THICKNESS = Thickness,
         MAXWIDTH = Width, LENGTH = Length, COREPLATFORM = Core_platfform, SCARCOUNT = N_of_Scars, COREABANDON = Core_abandon,
         RETOUCHEDPIECETYPOLOGY = Type)  # Renaming column

solutrean_tech <- solutrean_WIP %>%
  dplyr::select(ID, CLASS, LEVEL, UNIT, COREPREPPROD, PIECECOMPLETENESS, CORTEX, CORTEXLOCATION, PLATFORMTYPE,
                SCARCOUNT, THICKNESS, MAXWIDTH, LENGTH, COREPLATFORM, COREABANDON, RETOUCHEDPIECETYPOLOGY)

solutrean_points <- solutrean_points %>% # Creating ID from 
  mutate_at("YEAR", str_replace, "20", "") %>% # Removing initial part of the year to match IDs
  tidyr::unite(col='ID', all_of(cols_to_concat_shelter), sep="-", remove=TRUE)

solutrean_tech <- bind_rows(solutrean_tech, solutrean_points) # Uniting technology dataset with points (2167 to 2228)

solutrean_tech <- solutrean_tech %>% # Originally 2228
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets (down to 1993)

shelter_RM_WIP <- shelter_RM %>%
  mutate_at("ID", str_replace, "VB", "") %>% # Remove "VB" from ID
  mutate(RMTYPE = case_when(
    RMTYPE == "Type 2" ~ "T2",
    RMTYPE == "Chalcedony" ~ "T1", # Altering first group names to final group names. Subtypes are maintained as before.
    RMTYPE == "TYPE 15" ~ "T3",
    RMTYPE == "TYPE 7" ~ "T4",
    RMTYPE == "TYPE 12" ~ "T5",
    RMTYPE == "TYPE 11" ~ "T6",
    RMTYPE == "TYPE 13" ~ "T7",
    RMTYPE == "TYPE 11F" ~ "T8",
    RMTYPE == "TYPE 13B" ~ "T9",
    RMTYPE == "Oolitic" ~ "T10",
    RMTYPE == "TYPE GB" ~ "T11",
    RMTYPE == "ND_A" ~ "INDET",
    .default = RMTYPE)) %>%
  select(AREA, ID, RMTYPE, SUBTYPE, WEIGHT, CORTEXTHICKNESS, CORTEXAPPEARANCE) %>%  # Select variables
  distinct(ID, .keep_all = TRUE) # Keeping individual IDs to avoid buckets

solutrean_dataset <- full_join(solutrean_tech, shelter_RM_WIP, by = "ID") # The complete dataset has 2818 (44 points) entries
solutrean_dataset <- solutrean_dataset %>%
  drop_na(RMTYPE) %>%   # Dropping NAs on the RM database drops dataset to 1514
  drop_na(CLASS) # Dropping NAs on the technological class drops dataset to 914

```

# Blank-to-core ratios

```{r}
#| label: ratio-blankcore

# Solutrean ratio
solutrean_ratiobc <- solutrean_dataset %>% 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" | CLASS == "Thinning flake" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(RMTYPE != "INDET") %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Blank/Core" = Blank / Core)

# Proto-Solutrean ratio
proto_ratiobc <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>%
  mutate("Blank/Core" = Blank / Core)

# Gravettian ratio
gravettian_ratiobc <- phase_dataset %>% 
  filter(Level %in% c(6, 7)) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "Core")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Blank/Core" = Blank / Core)

```

```{r}
#| label: tbl-bcratio
#| tbl-cap: Blank to core ratio by raw material of the Solutrean (a), Proto-Solutrean (b) and Gravettian (c) assemblages.
#| tbl-subcap: 
#|  - "Blank to core ratio by raw material of the Solutrean assemblage (levels A to C) of the Shelter area."
#|  - "Blank to core ratio by raw material of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#|  - "Blank to core ratio by raw material of the Gravettian assemblage (levels 6 and 7) of the Terrace area."
#| layout-nrow: 4

kable(head(solutrean_ratiobc))
kable(head(proto_ratiobc))
kable(head(gravettian_ratiobc))

```

# Tool-to-debitage ratios

```{r}
#| label: ratios-retouchdebitage

# Solutrean
solutrean_ratiotd <- solutrean_dataset %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "Flake" | CLASS == "ElongBlank" | CLASS == "Thinning flake" ~ "Blank",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    CLASS == "Retouched piece" ~ "RetouchedPiece",
    CLASS == "Retouched piece fragment" ~ "RetouchedPieceFrag",
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    CLASS == "RetouchedPieceFrag" & PIECECOMPLETENESS == "Proximal" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "RetouchedPiece")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Tool/Blank" = RetouchedPiece / Blank)

# Proto-Solutrean
proto_ratiotd <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    CLASS == "RetouchedPieceFrag" & PIECECOMPLETENESS == "Proximal" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "RetouchedPiece")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Tool/Blank" = RetouchedPiece / Blank)

proto5_ratiotd <- phase_dataset %>% 
  filter(Level == 5) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    CLASS == "RetouchedPieceFrag" & PIECECOMPLETENESS == "Proximal" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "RetouchedPiece")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Tool/Blank" = RetouchedPiece / Blank)

# Gravettian
gravettian_ratiotd <- phase_dataset %>% 
  filter(Level %in% c(6, 7)) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "Blank", # To add proximal fragments to ratio
    CLASS == "RetouchedPieceFrag" & PIECECOMPLETENESS == "Proximal" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  filter(CLASS %in% c("Blank", "RetouchedPiece")) %>% 
  count(RMTYPE, CLASS, sort = TRUE) %>% 
  group_by(RMTYPE, CLASS) %>%
  summarise(n=sum(n)) %>% 
  spread(CLASS, n, fill=0) %>% 
  mutate(sum = rowSums(across(where(is.numeric)))) %>% 
  filter(!sum <= 5) %>% 
  mutate("Tool/Blank" = RetouchedPiece / Blank)

```

```{r}
#| label: tbl-tdratio-solutrean
#| tbl-cap: Tools to debitage ratio by raw material of the Solutrean (a), Proto-Solutrean (b) and Gravettian (c) assemblages.
#| tbl-subcap: 
#|  - "Tools to debitage ratio by raw material of the Solutrean assemblage (levels A to C) of the Shelter area."
#|  - "Tools to debitage ratio by raw material of the Proto-Solutrean assemblage (levels 4E and top of level 5) of the Terrace area."
#|  - "Tools to debitage ratio by raw material of the Gravettian assemblage (levels 6 and 7) of the Terrace area."
#| layout-nrow: 4

kable(head(solutrean_ratiotd))
kable(head(proto_ratiotd))
kable(head(gravettian_ratiotd))

```

# Measurements

```{r}
#| label: stats-setup

# Solutrean statistics
# DS preparation
solutrean_stats <- solutrean_dataset %>% 
  filter(RMTYPE != "INDET") %>%
   mutate(CLASS = case_when(
    CLASS == "Thinning flake" ~ "ThinningFlake",
    CLASS == "ElongBlank" ~ "ElongatedProd",
    CLASS == "FlakeFrag" | CLASS == "ElongBlankFrag" ~ "BlankFrag",
    CLASS == "Retouched piece" ~ "RetouchedPiece",
    TRUE ~ CLASS)) %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS))

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
solutrean_stats_flakes <- solutrean_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
solutrean_stats_elong <- solutrean_stats %>% 
  filter(CLASS == "ElongatedProd") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
solutrean_stats_cores <- solutrean_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
solutrean_stats_retouch<- solutrean_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

# Proto-Solutrean statistics
# DS prep
proto_stats <- phase_dataset %>% 
  filter(Level %in% c("4E", "4E/5")) %>% 
  filter(RMTYPE != "INDET") %>% 
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>%
  mutate(elong = MAXWIDTH*2-LENGTH) %>%
  mutate(CLASS = case_when(
    CLASS == "Blank" & elong > 0 ~ "Flake", # QUESTION: do I keep flakes and elong separate or join into blanks?
    CLASS == "Blank" & elong <= 0 ~ "ElongatedProd",
    TRUE ~ CLASS))

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
proto_stats_flakes <- proto_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
proto_stats_elong <- proto_stats %>% 
  filter(CLASS == "ElongatedProd") %>% 
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
proto_stats_cores <- proto_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
proto_stats_retouch <- proto_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

# Gravettian statistics
# DS prep
gravettian_stats <- phase_dataset %>%
  filter(Level %in% c(6, 7)) %>%
  filter(RMTYPE != "INDET") %>%
  mutate(CLASS = case_when(
    CLASS == "BlankFrag" & PIECECOMPLETENESS == "Proximal" ~ "BlankFragProx", # To add proximal fragments to ratio
    TRUE ~ CLASS)) %>% 
  mutate(elong = MAXWIDTH*2-LENGTH) %>%
  mutate(CLASS = case_when(
    CLASS == "Blank" & elong > 0 ~ "Flake", # QUESTION: do I keep flakes and elong separate or join into blanks?
    CLASS == "Blank" & elong <= 0 ~ "ElongatedProd",
    TRUE ~ CLASS))

# Tables with individual class. For supplementary information if necessary
# For flakes (complete)
gravettian_stats_flakes <- gravettian_stats %>% 
  filter(CLASS == "Flake") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For elongated blanks (complete)
gravettian_stats_elong <- gravettian_stats %>% 
  filter(CLASS == "ElongatedProd") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For cores
gravettian_stats_cores <- gravettian_stats %>% 
  filter(CLASS == "Core") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))
# For retouched
gravettian_stats_retouch <- gravettian_stats %>% 
  filter(CLASS == "RetouchedPiece") %>%
  select(RMTYPE, CLASS, THICKNESS, MAXWIDTH, LENGTH, WEIGHT) %>% 
  group_by(RMTYPE, CLASS) %>% 
  summarise(across(c(THICKNESS, MAXWIDTH, LENGTH, WEIGHT), list(mean = mean, sd = sd, max = max, min = min)))

```

```{r}
#| label: tbl-solutrean
#| tbl-cap: "Stats for Solutrean occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(solutrean_stats_flakes))
kable(head(solutrean_stats_elong))
kable(head(solutrean_stats_cores))
kable(head(solutrean_stats_retouch))

```

```{r}
#| label: tbl-proto
#| tbl-cap: "Stats for Proto-Solutrean occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(proto_stats_flakes))
kable(head(proto_stats_elong))
kable(head(proto_stats_cores))
kable(head(proto_stats_retouch))

```

```{r}
#| label: tbl-gravettian
#| tbl-cap: "Stats for Gravettian occupation."
#| tbl-subcap: 
#|  - "Stats for flakes"
#|  - "Stats for elongated blanks"
#|  - "Stats for cores"
#|  - "Stats for retouched"
#| layout-nrow: 4

kable(head(gravettian_stats_flakes))
kable(head(gravettian_stats_elong))
kable(head(gravettian_stats_cores))
kable(head(gravettian_stats_retouch))

```